Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-25 paper.qmd
Update latex

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography:
  # cruse.bib
  - bibliography.bib
  - cruse.bib
  - references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  docx: default
# # # # Uncomment for formatted version
#   elsevier-pdf:
#     linestretch: 2
#     keep-tex: true
#     journal:
#       name: European Transport Research Review
#       formatting: preprint
#       model: 3p
#       cite-style: super
# # # #   # pdf: default

@@

```{r}
#| label: tbl-tools
# For PDF:
# knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.", booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE) |>
#   kableExtra::column_spec(1, width = "13em") |>
#   kableExtra::column_spec(2, width = "15em") |>
#   kableExtra::column_spec(3, width = "12em") |>
#   kableExtra::column_spec(4, width = "5em")
# For HTML:
knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.")
```

>>>>

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography:
  # cruse.bib
  - bibliography.bib
  - cruse.bib
  - references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  # docx: default
# # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default

@@

```{r}
#| label: tbl-tools
# For PDF:
knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.", booktabs = TRUE) |>
  kableExtra::kable_styling(latex_options = "striped", full_width = FALSE) |>
  kableExtra::column_spec(1, width = "13em") |>
  kableExtra::column_spec(2, width = "15em") |>
  kableExtra::column_spec(3, width = "12em") |>
  kableExtra::column_spec(4, width = "5em")
# For HTML:
# knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.")
```

### 2024-03-25 paper.tex
Update latex

Another motivation for cycle network planning at the European level is
the Road Infrastructure Safety Management (RISM) directive (2008/96/EC),
which requires member states to implement a road safety management
system (RSMS) for all public roads. Specifically, ``Member States shall
ensure that the ranking of high accident concentration sections and the
network safety ranking are carried out'' \citep{directiv2008}. Given
that `safety' in this context is usefully quantified as the number of
people killed or seriously injured (KSI) per distance traveled, the
directive requires estimation of distance traveled by mode, down to the
road link level. For active modes, about which there is a paucity of
data compared with motorized modes, this is a major challenge. Better
data to inform road safety policies and interventions is a motivation
for better estimates of `baseline' levels of physical activity at high
levels of geographic resolution \citep{tait2023}.

National governments are increasingly acting on the evidence. In
Ireland, the Road Safety Authority (RSA) has set the target of halving
the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key travel corridors to be
identified `cycle proofed'. Cycling in Ireland represents only 3\% of
total modal share as of the 2016 Census, but accounts for 20\% of
serious injuries and 7\% of all fatalities, driving the development of
cycling policy and infrastructure plans in Ireland. The main frameworks
underpinning these efforts are the Climate Action Plan
\citep{climate2022a}, the National Development Plan, and the National
Roads 2040 strategy \citep{national2023}. These policy frameworks are
backed-up by public support for active travel interventions: while only
3\% of trips in Ireland are made by cycling according to the latest
Census data from 2016 ``83\% of respondents supported enhanced walking
and cycling facilities'' \citep{sustaina2019}.

At the regional level within Ireland, the recently published Greater
Dublin Area Transport Strategy \citep{greater} reinforces these findings
of high support for and capability for cycling: ``nearly a quarter of
adults cycle at least once a week in the Dublin Metropolitan Area'' with
cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around
40\% of the population living in Dublin, suggests that around 150,000
cars could be removed nationwide just by achieving Dublin levels of
cycling in all counties (notwithstanding existing cycling trips and
differences in trip distances). 7 in 10 trips in Ireland are by car
\citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as ebikes extend the range of this
mode'' \citep{greater}.

@@

\begin{quote}
Direct routes with short journey times were found to be the most
important positive variable for existing cyclists and non-cyclists in
determining route choice. This is followed by infrastructure type, the
number of junctions along the route, traffic speed and cyclist volumes.
In terms if infrastructure, regardless of the level of cycling
confidence, routes which have `no facilities' or `bus/cycle lanes' are
the least favoured cycle route types.
\end{quote}

@@

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratize
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarized
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarizing nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel school \citep{goodman2019}
  as covered in previously published research on open access national
  cycle network planning tools --- the results capture a higher
  proportion of cycling potential including key rural trips which are
  often under-represented in models of active travel.
\end{itemize}

>>>>

Another motivation for cycle network planning at the European level is
the Road Infrastructure Safety Management (RISM) directive (2008/96/EC),
which requires member states to implement a road safety management
system (RSMS) for all public roads. Specifically, ``Member States shall
ensure that the ranking of high accident concentration sections and the
network safety ranking are carried out'' \citep{directiv2008}. Given
that `safety' in this context is usefully quantified as the number of
people killed or seriously injured (KSI) per distance traveled, the
directive requires estimation of distance traveled by mode, down to the
road link level. For active modes, about which there is a paucity of
data compared with motorized modes, this is a major challenge. Better
data to inform road safety policies and interventions is a motivation
for better estimates of `baseline' levels of physical activity at high
geographic resolutions \citep{tait2023}.

National governments are increasingly acting on the evidence. In
Ireland, the Road Safety Authority (RSA) has set the target of halving
the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key travel corridors to be
identified and `cycle proofed'. Cycling in Ireland represents only 3\%
of total modal share as of the 2016 Census, but accounts for 20\% of
serious injuries and 7\% of all fatalities. Poor perception of safety
has been found to represent the most important barrier to increased
cycling in Ireland\citep{brick2018}, and the need to improve road safety
drives the development of national cycling policy and infrastructure
plans. The main frameworks underpinning these efforts are the Climate
Action Plan \citep{climate2022a}, the National Development Plan, and the
National Roads 2040 strategy \citep{national2023}.

At the regional level within Ireland, the recently published Greater
Dublin Area Transport Strategy \citep{greater} reveals the high support
for and capability for cycling: ``nearly a quarter of adults cycle at
least once a week in the Dublin Metropolitan Area'' with cycling in the
Dublin area taking up to 60,000 cars off the road today. Extrapolating
this on a per population basis across Ireland, with around 40\% of the
population living in Dublin, suggests that around 150,000 cars could be
removed nationwide just by achieving Dublin levels of cycling in all
counties (notwithstanding existing cycling trips and differences in trip
distances). 7 in 10 trips in Ireland are by car \citep{national} .
Cycling has the potential to replace a large proportion of these trips:
``A high priority must also be given to cyclists, because trips by this
mode have the potential to replace trips by private car, most
specifically for short to medium distance trips, but increasingly for
longer trips as ebikes extend the range of this mode'' \citep{greater}.

@@

\begin{quote}
Direct routes with short journey times were found to be the most
important positive variable for existing cyclists and non-cyclists in
determining route choice. This is followed by infrastructure type, the
number of junctions along the route, traffic speed and cyclist volumes.
In terms of infrastructure, regardless of the level of cycling
confidence, routes which have `no facilities' or `bus/cycle lanes' are
the least favoured cycle route types.
\end{quote}

@@

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratize
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarized
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarizing nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel to school
  \citep{goodman2019} as covered in previously published research on
  open access national cycle network planning tools --- the results
  capture a higher proportion of cycling potential including key rural
  trips which are often under-represented in models of active travel.
\end{itemize}
